The dynamics of starvation and recovery

### 2017-06-15 Manuscript/NatureComm/YKR_NatureComm.tex
final changes for naturecomm first submission

\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
%\usepackage[pdftex]{graphicx}

@@

%\usepackage{epstopdf}
%\usepackage{array}
%\usepackage{tabularx}
%\usepackage{multirow}
\usepackage{color}
%\usepackage{multibox}
%\usepackage{rotating}
%\usepackage{lineno}
%\usepackage[left]{lineno}
\usepackage[comma,sort&compress]{natbib}
\bibpunct{}{}{,}{s}{}{;}
%\usepackage{authblk}
%\usepackage{multicol}

@@

\usepackage{bibunits}

%\linenumbers
%\setlength\linenumbersep{3pt}

@@

\begin{bibunit}[unsrt]

@@

% \noindent {\bf Role of allometry} \\
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
While there are no a priori constraints on the parameters in the NSM, most
organisms correspond to restricted portions of the parameter space.  We
use allometric scaling relations to constrain the covariation of rates in a
principled and biologically meaningful manner (see Methods).  Allometric scaling relations
highlight common constraints and average trends across large ranges in body
size and species diversity. Many of these relations can be derived from a
small set of assumptions and in the Methods we describe our framework to determine the
covariation of timescales and rates across a range of body sizes for each of
the key parameters of our model (cf. ref.~\citep{Yodzis:1992hg}).  
% We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.

Nearly all of the rates described in the NSM are determined by consumer
metabolism, which can be used to describe a variety of organismal features
\citep{Brown:2004wq}.
We derive relationships for the rates of reproduction, starvation, recovery, and mortality based on first principles, and as a function of an organism's body size and metabolic rate (see Methods).
Because we aim to explore the starvation-recovery dynamics as a function of an organism's body mass $M$, we parameterize these rates in terms of the \emph{percent} gain and loss of the asymptotic (maximum) body mass, $\epsilon M$, where different values of $\epsilon$ defines different states of the consumer (Fig.~\ref{fig:growth}).
Although the rate equations \eqref{eq:system} are general and can in principle be used to explore the starvation recovery dynamics for most organisms, here we focus on allometric relationships for terrestrial-bound lower trophic level endotherms (see Supplementary Information for values), specifically herbivorous mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$. Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.\\

@@

We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}a-c, white points).
These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig. \ref{fig:ext}a-c), suggesting that the population dynamics of larger organisms are more sensitive to smaller changes in physiological rates controlling starvation and recovery.
% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\

@@

\putbib[aa_starving3]

\end{bibunit}

@@

%%%%%%%%%%%%%%%%%%%%%%%%
% SUPPLEMENTARY MATERIAL
%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage

\begin{bibunit}[unsrt]
{\bf Supporting Information for ``The dynamics of starvation and recovery''}\\

{\bf Mechanisms of Starvation and Recovery}
Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce. For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs. \citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective). This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g. \citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}. Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).

We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
\begin{align}
\begin{split}
\dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
& \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
\end{split}
\end{align}
where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
%(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)

%\begin{align}
%\begin{split}
%\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
%\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
%\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
%&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
%\end{split}
%\end{align}
%
%In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability. Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources. The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
%A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
%
%Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification. As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$. Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to

%where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.

We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
%(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
\begin{align}
\begin{split}
\dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
\dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
\dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
\end{split}
\end{align}
If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
\begin{align}
\begin{split}
\dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
\dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
\dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
\end{split}
\end{align}
where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$. The above equations represent the system of equations presented in the main text.
\\

{\bf Parameter Values and Estimates}
All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies. Below we outline previous measurements and simple estimates of the parameters.

Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs. \citep{West:2001bv,moses2008rmo} and the supplement for Ref. \citep{hou}). We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends. The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.  For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs. \citep{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate. We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).

An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction. This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref. \citep{peters1986ecological} for a review). We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.

The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g. Refs. \citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}. The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}. For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.

For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}. We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range. Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.

%We also connect the resource growth rate to the total metabolic rate of an organism. That is, we are interested in the relative rates of resource recovery and consumption by the total population. From \citep{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$). Considering an energy density of $18200$ (J g$^{-1}$) of grass \citep{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$). This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.

Typically the value of $\xi=C/k$ should roughly be $2$. The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext. Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages. Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
\begin{equation}
B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
\begin{equation}
B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
\end{equation}
Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.

>>>>

\RequirePackage{lineno}
\documentclass[twocolumn,preprintnumbers,amsmath,amssymb,superscriptaddress]{revtex4}
%\usepackage[pdftex]{graphicx}

@@

%\usepackage{epstopdf}
%\usepackage{array}
%\usepackage{tabularx}
%\usepackage{multirow}
\usepackage{color}
%\usepackage{multibox}
%\usepackage{rotating}
\usepackage{lineno}
%\usepackage[left]{lineno}
\usepackage[comma,sort&compress]{natbib}
\bibpunct{}{}{,}{s}{}{;}
%\usepackage{authblk}
%\usepackage{multicol}

@@

% \usepackage{bibunits}

\linenumbers
\setlength\linenumbersep{3pt}

@@

% \begin{bibunit}[unsrt]

@@

% \noindent {\bf Role of allometry} \\
%[Link Allometry stuff to our model - what does it provide to the story]
%[Introduce biological and linked constraints]
%[Allometries capture vast amounts of diversity via a single parameter --- body size]
%[Allometries have captured intterest etc across multiple scales and biological classes of organisms]
While there are no {\it a priori} constraints on the parameters in the NSM, we expect that each species should be restricted to a distinct portion of the parameter space.  We
use allometric scaling relations to constrain the covariation of rates in a
principled and biologically meaningful manner (see Methods).  Allometric scaling relations
highlight common constraints and average trends across large ranges in body
size and species diversity. Many of these relations can be derived from a
small set of assumptions and in the Methods we describe our framework to determine the
covariation of timescales and rates across a range of body sizes for each of
the key parameters of our model (cf. ref.~\citep{Yodzis:1992hg}).  
% We are thereby able to define the regime of dynamics occupied by the entire class of mammals, along with the key differences between the largest and smallest mammals.

Nearly all of the rates described in the NSM are determined by consumer
metabolism, which can be used to describe a variety of organismal features
\citep{Brown:2004wq}.
We derive relationships for the rates of reproduction, starvation, recovery, and mortality based on first principles, and as a function of an organism's body size and metabolic rate (see Methods).
Because we aim to explore the starvation-recovery dynamics as a function of an organism's body mass $M$, we parameterize these rates in terms of the \emph{percent} gain and loss of the asymptotic (maximum) body mass, $\epsilon M$, where different values of $\epsilon$ define different states of the consumer (Fig.~\ref{fig:growth}; see Methods for derivations of allometrically constrained rate equations).
Although the rate equations \eqref{eq:system} are general and can in principle be used to explore the starvation recovery dynamics for most organisms, here we focus on allometric relationships for terrestrial-bound lower trophic level endotherms (see Supplementary Information for values), specifically herbivorous mammals, which range from a minimum of $M\approx1$g (the Etruscan shrew \emph{Suncus etruscus}) to a maximum of $M\approx10^7$g (the early Oligocene Indricotheriinae and the Miocene Deinotheriinae).
Investigating other classes of organisms would simply involve altering the metabolic exponents and scalings associate with $\epsilon$. Moreover, we emphasize that our allometric equations describe mean relationships, and do not account for the (sometimes considerable) variance associated with individual species.\\

@@

We find that the allometrically constrained values of $\sigma$ and $\rho$ fall squarely within the extinction refuge across a range of $M$ (Fig. \ref{fig:ext}a-c, white points).
These values are close enough to the Hopf bifurcation to avoid low steady-state densities, and far enough away to avoid large-amplitude transient cycles.
The feature that allometric values of $\sigma$ and $\rho$ fall within this relatively small window supports the possibility that a selective mechanism has constrained the physiological conditions that drive starvation and recovery rates within populations.
Such a mechanism would select for organism physiology that generates appropriate $\sigma$ and $\rho$ values that serve to minimize extinction risk.
This selection could occur via the tuning of body fat percentages, metabolic rates, and biomass maintenance efficiencies.
We also find that as body size increases, the amount of low extinction risk parameter space becomes smaller (Fig. \ref{fig:ext}a-c), suggesting that the population dynamics of larger organisms are more sensitive to variability in physiological rates controlling starvation and recovery.
This finding is in accordance with, and may serve as contributing support for, observations of increased extinction risk among larger mammals \cite{Liow:2008jx}.
% OLD WORDING Such a mechanism would involve a feedback between the dynamics of the population and the fitness of individuals within the population, though to what extent the dynamics of the population influence rates of starvation and recovery would also involve potential tradeoffs in reproduction and somatic maintenance.
To summarize, our finding that the allometrically-determined parameters fall within this low extinction probability region suggests that the NSM dynamics may both drive---and constrain---natural animal populations.\\

@@

\bibliography{aa_starving3}

% \end{bibunit}

@@

%%%%%%%%%%%%%%%%%%%%%%%%
% SUPPLEMENTARY MATERIAL
%%%%%%%%%%%%%%%%%%%%%%%%
% 
% 
% \clearpage
% 
% \begin{bibunit}[unsrt]
% {\bf Supporting Information for ``The dynamics of starvation and recovery''}\\
% 
% {\bf Mechanisms of Starvation and Recovery}
% Our overall goal is to understand the dynamics of starvation, recovery, reproduction, and resource competition, where our framework partitions starvation and reproduction into two classes of the consumer: a full class that is able to reproduce and a hungry class that experiences mortality at a given rate and is unable to reproduce. For the dynamics of growth, reproduction, and resource consumption, past efforts have combined the overall metabolic rate as dictated by body size with a growth rate that is dependent on resource abundance and, in turn, dictates resource consumption (see Refs. \citep{Kempes:2012hy,kempes2014morphological} for a brief review of this perspective). This approach has been used to understand a range of phenomena including a derivation of ontogenetic growth curves from a partitioning of metabolism into maintenance and biosynthesis (e.g. \citep{West:2001bv,moses2008rmo,hou,Kempes:2012hy}) and predictions for the steady-state resource abundance in communities of cells \citep{kempes2014morphological}. Here we leverage these mechanisms, combined with several additional concepts, to define our Nutritional State Model (NSM).
% 
% We consider the following generalized set of explicit dynamics for starvation, recovery, reproduction, and resource growth and consumption
% \begin{align}
% \begin{split}
% \dot{F_{d}} &= \lambda_{max} F_{d} + \rho_{max}R_{d}H_{d}/k - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
% \dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho_{max}R_{d} H_{d}/k - \mu H_{d},  \\
% \dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) -\\
% & \left[\left(\frac{\rho_{max}R_{d}}{Y_{H}k}+P_{H}\right)H_{d}+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)F_{d}\right].
% \end{split}
% \end{align}
% where each term has a mechanistic meaning that we detail below (we will denote the dimensional equations with $_{d}$ before introducing the non-dimensional form that is presented in the main text). In the above equations $Y$ represents the yield coefficient (e.g., Refs. \citep{pirt,Heijnen}) which is the quantity of resources required to build a unit of organism (gram of mammal produced per gram of resource consumed) and $P$ is the specific maintenance rate of resource consumption (g resource $\cdot$ s$^{-1}$ $\cdot$ g organism). If we pick $F_{d}$ and $H_{d}$ to have units of (g organisms $\cdot$ m$^{-2}$), then all of the terms of $\dot{R_{d}}$, such as $\frac{\rho\left(R_{d}\right)}{Y}H_{d}$, have units of (g resource $\cdot$ m$^{-2}$ $\cdot$ s$^{-1}$) which are the units of net primary productivity (NPP), a natural choice for $\dot{R_{d}}$. This choice also gives $R_{d}$ as (g $\cdot$ m$^{-2}$) which is also a natural unit and is simply the biomass density. In this system of units $\alpha$ (s$^{-1}$) is the specific growth rate of $R_{d}$, and $C$ is the carrying capacity, or maximum density, of $R_{d}$ in a particular environment, and $k$ is the half-saturation constant (half the density of resources that would lead to maximum growth).
% %(note this is not fully explicit because I don't know how to deal with the response of $\sigma$ to resources, although I have an idea for a derivation which may be necessary given the following approximations)
% 
% 
% 
% 
% %\begin{align}
% %\begin{split}
% %\dot{F_{d}} &= \lambda\left(R_{d}\right) F_{d} + \rho\left(R_{d}\right)H_{d} - \sigma \left(1-\frac{R_{d}}{C}\right)F_{d},  \\
% %\dot{H_{d}} &= \sigma \left(1-\frac{R_{d}}{C}\right)F_{d} - \rho\left(R_{d}\right)H_{d} - \mu H_{d},  \\
% %\dot{R_{d}} &= \alpha R_{d}\left(1-\frac{R_{d}}{C}\right) - \\
% %&\left[\left(\frac{\rho\left(R_{d}\right)}{Y}+P_{H}\right)H_{d}+\left(\frac{\lambda\left(R_{d}\right)}{Y}+P_{F}\right)F_{d}\right],
% %\end{split}
% %\end{align}
% %
% %In this set of equations $\lambda\left(R_{d}\right)$ and $\rho\left(R_{d}\right)$ are the growth and recovery rates as functions of the current resource availability. Typically these can be written as $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ or $\lambda\left(R_{d}\right)=\lambda_{max}S\left(R_{d}\right)$ where $\lambda_{max}$ and $\rho_{max}$ are the maximum growth and recovery rates respectively, which scale with body size as discussed later, and $S\left(R_{d}\right)$ is a saturating function of resources. The saturating function could, for example, be a Michaelis-Menten or Monod function of the form $\frac{R_{d}}{k+R_{d}}$, where $k$ is the half-saturation constant.
% %A simplified version of the Michaelis-Menten or Monod functional form, which captures the essential features, is a linear function that saturates to a constant value above a certain abundance of $R_{d}$.
% %
% %Before describing the values of each of these constants, and a general non-dimensionalization of the system of equations, it is important to consider the resource regimes associated with the above equations which lead to a simplification. As discussed above, the resource saturation function should be defined by a linear regime proportional to $R_{d}$ when $R_{d}<<k$, and a constant value for $R_{d}>>k$. Thus for hungry individuals, $H_{d}$, where $R_{d}<<k$, we have that $\rho\left(R_{d}\right)\approx\rho_{max}R_{d}/k$, and for the full class, $F_{d}$, of organisms $\lambda\left(R_{d}\right)\approx\lambda_{max}$, such that the above relationships reduce to
% 
% 
% 
% 
% 
% %where $\beta=\frac{\lambda_{max}}{Y_{F}}+P$ which is just a constant that depends on the size of an organisms via the allometries for $\lambda_{max}$ and $P$ discussed later.
% 
% We can formally non-dimensionalize this system by choosing the general rescaling of $F=fF_{d}$, $H=fH_{d}$, $R=qR_{d}$, $t=st_{d}$, in which case our system of equations becomes
% %(ignoring the $\sigma (1-R)F$ terms which I don't have a dimensional form for yet),:
% \begin{align}
% \begin{split}
% \dot{F} &= \frac{1}{s}\left[\lambda_{max} F + \rho_{max}\frac{R}{qk}H - \sigma \left(1-\frac{R}{qC}\right)F\right],  \\
% \dot{H} &= \frac{1}{s}\left[\sigma \left(1-\frac{R}{qC}\right)F - \rho_{max}\frac{R}{qk} H - \mu H\right],  \\
% \dot{R} &= \frac{1}{s}\left[\alpha R\left(1-\frac{R}{qC}\right) -\frac{q}{f}\left[\left(\frac{\rho_{max}R}{Y_{H}kq}+P_{H}\right)H+\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right) F\right]\right].
% \end{split}
% \end{align}
% If we make the natural choice of $s=1$, $q=1/C$, and $f=1/Y_{H}k$, then we are left with
% \begin{align}
% \begin{split}
% \dot{F} &= \lambda F + \xi \rho RH - \sigma \left(1-R\right)F,  \\
% \dot{H} &= \sigma \left(1-R\right)F - \xi \rho RH - \mu H,  \\
% \dot{R} &= \alpha R\left(1-R\right) -\left(\rho R+\delta\right)H-\beta F
% \end{split}
% \end{align}
% where we have dropped the subscripts on $\lambda_{max}$ and $\rho_{max}$ for simplicity, and $\xi=C/k$, $\delta=Y_{H}kP_{H}/C$, and $\beta=Y_{H}k\left(\frac{\lambda_{max}}{Y_{F}}+P_{F}\right)/C$. The above equations represent the system of equations presented in the main text.
% \\
% 
% {\bf Parameter Values and Estimates}
% All of the parameter values employed in our model have either been directly measured in previous studies or can be estimated from combining several previous studies. Below we outline previous measurements and simple estimates of the parameters.
% 
% Metabolic rate has been generally reported to follow an exponent close to $\eta=0.75$ (e.g., Refs. \citep{West:2001bv,moses2008rmo} and the supplement for Ref. \citep{hou}). We make this assumption in the current paper, although alternate exponents, which are know to vary between roughly $0.25$ and $1.5$ for single species \citep{moses2008rmo}, could be easily incorporated into our framework, and this variation is effectively handled by the $20\%$ variations that we consider around mean trends. The exponent not only defines several scalings in our framework, but also the value of the metabolic normalization constant, $B_{0}$, given a set of data.  For mammals the metabolic normalization constant has been reported to vary between $0.018$ (W g$^{-0.75}$) and $0.047$ (W g$^{-0.75}$; Refs. \citep{hou,West:2001bv}, where the former value represents basal metabolic rate and the latter represents the field metabolic rate. We employ the field metabolic rate for our NSM model which is appropriate for active mammals (Table 1).
% 
% An important feature of our framework is the starting size, $m_{0}$, of a mammal which adjusts the overall timescales for reproduction. This starting size is known to follow an allometric relationship with adult mass of the form $m_{0}=n_{0}M^{\upsilon}$ where estimates for the exponent range between $0.71$ and $0.94$ (see Ref. \citep{peters1986ecological} for a review). We use $m_{0}=0.097M^{0.92}$ \citep{blueweiss1978relationships} which encompasses the widest range of body sizes \citep{peters1986ecological}.
% 
% The energy to synthesize a unit of biomass, $E_{m}$, has been reported to vary between $1800$ to $9500$ (J g$^{-1}$) (e.g. Refs. \citep{West:2001bv,moses2008rmo,hou}) in mammals with a mean value across many taxonomic groups of $5,774$ (J g$^{-1}$) \citep{moses2008rmo}. The unit energy available during starvation, $E^{\prime}$, could range between $7000$ (J g$^{-1}$), the return of the total energy stored during ontogeny \citep{hou} to a biochemical upper bound of $E^{\prime}=36,000$ (J g$^{-1}$) for the energetics of palmitate \citep{stryer,hou}. For our calculations we use the measured value for bulk tissues of $7000$ which assumes that the energy stored during ontogeny is returned during starvation \citep{hou}.
% 
% For the scaling of body composition it has been shown that fat mass follows $M_{\rm fat}=f_{0}M^{\gamma}$, with measured  relationships following  $0.018M^{1.25}$ \citep{Dunbrack:1993ec}, $0.02M^{1.19}$ \citep{Lindstedt:1985hm}, and $0.026M^{1.14}$ \citep{Lindstedt:2002td}. We use the values from \citep{Lindstedt:1985hm} which falls in the middle of this range. Similarly, the muscle mass follows $M_{\rm musc}=u_{0}M^{\zeta}$ with $u_{0}=0.383$ and $\zeta=1.00$ \citep{Lindstedt:2002td}.
% 
% %We also connect the resource growth rate to the total metabolic rate of an organism. That is, we are interested in the relative rates of resource recovery and consumption by the total population. From \citep{allen2002global} the total resource use of a population with an individual body size of $M$ is given by $B_{pop}=0.00061x^{-0.03}$ (W m$^{-2}$). Considering an energy density of $18200$ (J g$^{-1}$) of grass \citep{estermann} and an NPP between and $1.59\times10^{-6}$ and $7.92\times10^{-5}$ (g s$^{-1}$ m$^{-2}$) would give a range of resource rates between  $0.029$ and $1.44$ (W m$^{-2}$). This gives a ratio of total resource consumption to supply rates between $0.00042$ and $0.021$, and we used a value of $0.002$ in our calculations and simulations.
% 
% Typically the value of $\xi=C/k$ should roughly be $2$. The value of $\rho$, $\lambda$, $\sigma$, and $\mu$ are all simple rates (note that we have not rescaled time in our non-dimensionalization) as defined in the maintext. Given that our model considers transitions over entire stages of ontogeny or nutritional states, the value of $Y$ must represent yields integrated over entire life stages. Given an energy density of $E_{d}=18200$ (J g$^{-1}$) for grass \citep{estermann} the maintenance value is given by $P_{F}=B_{0}M^{3/4}/ME_{d}$, and the yield for a full organism will be given by $Y_{F}=ME_{d}/B_{\lambda}$ (g individual $\cdot$ g grass $^{-1}$), where $B_{\lambda}$ is the lifetime energy use for reaching maturity given by
% \begin{equation}
% B_{\lambda}=\int_{0}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
% \end{equation}
% Similarly, the maintenance for hungry individuals is $P_{H}=B_{0}(\epsilon_{\sigma}M)^{3/4}/(\epsilon_{\sigma}M)E_{d}$, and the yield for hungry individuals (representing the cost on resources to return to the full state) is given by $Y_{H}=ME_{d}/B_{\rho}$ where
% \begin{equation}
% B_{\rho}=\int_{\tau\left(\epsilon_{\sigma}\epsilon_{\lambda}\right)}^{t_{\lambda}}B_{0}m\left(t\right)^{\eta}dt.
% \end{equation}
% Taken together, these relationships allow us to calculate $\rho$, $\delta$, and $\beta$.
% 
